\section{Background}
\label{sec:background}

\subsection{DETR and its descendants}
DETR~\cite{carion2020detr} reframed detection as direct set prediction. A CNN backbone produces a feature map; a transformer encoder--decoder operates on a flattened token sequence; $N$ object queries cross-attend over encoder memory across $L$ decoder layers; a Hungarian matcher assigns predictions to ground-truth boxes for loss computation; and inference is NMS-free. Standard transformer attention reads
\begin{equation}
    \mathrm{Attn}(Q, K, V) = \mathrm{softmax}\!\left(\tfrac{QK^\top}{\sqrt{d_k}}\right) V,
    \label{eq:attn}
\end{equation}
with multi-head attention concatenating $h$ such operations. DETR's main weakness is slow convergence and weak small-object detection, attributed to the global, dense attention pattern over a single low-resolution feature map.

\subsection{Deformable attention and multi-scale memory}
Deformable DETR~\cite{zhu2021deformable} replaces global attention with a sparse mechanism: each query samples a small number of points $K$ from each of $L$ feature levels at learned offsets $\Delta p_{lqk}$ around a reference point $\hat p_q$:
\begin{equation}
    \mathrm{MSDA}(Q, \hat p_q, \{x^l\}) = \sum_{l=1}^{L} \sum_{k=1}^{K} A_{lqk}\, W_l\, x^l\!\bigl(\hat p_q + \Delta p_{lqk}\bigr),
    \label{eq:msda}
\end{equation}
where $A_{lqk}$ are learned attention weights that sum to one across $(l,k)$ for each query $q$. This made multi-scale memory tractable and dramatically sped up convergence.

\subsection{RT-DETR architecture}
RT-DETR~\cite{lyu2024rtdetr} is a real-time variant built on this lineage. Its hybrid encoder splits into two complementary blocks:
\begin{itemize}
    \item \textbf{AIFI} --- a transformer block applied only to the smallest feature map ($S_5$, stride 32), where each token has the largest receptive field and intra-scale interaction is most useful.
    \item \textbf{CCFF} --- a CNN-based feature fusion that combines features across pyramid levels (PAN-style), keeping cross-scale aggregation cheap.
\end{itemize}
The flattened multi-level token sequence (the \emph{encoder memory}) is then attended over by the decoder via Equation~\ref{eq:msda}. Queries are initialized from the top-$N$ encoder predictions (anchor-style, à la DAB-DETR~\cite{liu2022dabdetr}), with optional contrastive denoising auxiliary queries during training~\cite{zhang2023dino}.

For our R50-backbone variant with input $640\!\times\!640$ and the additional P2 (stride-4) feature level we adopt in Section~\ref{sec:method}, the memory has shape $[B,\, L,\, d{=}256]$ with
\begin{equation}
    L = \underbrace{160{\times}160}_{\text{P2}} + \underbrace{80{\times}80}_{\text{P3}} + \underbrace{40{\times}40}_{\text{P4}} + \underbrace{20{\times}20}_{\text{P5}} = 25600 + 6400 + 1600 + 400 = 34000.
\end{equation}
Of these, 25,600 tokens (75\%) live at stride 4 (P2), where each token covers $4\!\times\!4$ image pixels --- the natural resolution for small objects.

\subsection{Why small objects are hard for transformers}
Transformer detectors share a structural disadvantage on small objects relative to FPN-equipped CNN detectors:
\begin{enumerate}
    \item Each query receives information about all objects in an image through a fixed-capacity attention budget; small objects get little of that budget when the image is dominated by large objects.
    \item The encoder memory is computed once before any decoder layer fires, so refinement is one-way: queries can attend over memory, but memory has no path to be revised in light of what the queries discover.
    \item Most published transformer detectors only use stride-$\{8, 16, 32\}$ feature levels (P3--P5). Stride 4 (P2) is omitted for cost reasons --- but it is the level where a 16-pixel object is still 4x4 tokens, while at stride 8 it is 2x2 and at stride 32 it is sub-token.
\end{enumerate}
Our two architectural moves --- adding the P2 level and introducing a feedback path --- target points (3) and (2) respectively. Point (1) is left to the broader query-allocation literature.
