\section{Limitations}
\label{sec:limitations}

We acknowledge several limitations, in roughly decreasing order of how much they would change the conclusion:

\begin{enumerate}
    \item \textbf{Single seed.} All numbers are from one training run. A multi-seed sweep would tighten the confidence interval on $\Delta\mathrm{AP}_S$ and confirm whether $+0.99$ is reproducible.
    \item \textbf{Floor and mask not isolated.} v2 changes both fixes simultaneously, so we cannot say from this experiment whether one alone is sufficient. A $2{\times}2$ ablation is the obvious next step.
    \item \textbf{Below the published RT-DETR baseline.} v2 reaches $\mathrm{AP}_S = 34.25$ vs the published $34.7$. Our claim is the $+0.99$ causal contribution, not a new state of the art --- the absolute deficit is explained by the schedule (13-epoch finetune vs the paper's 6$\times$ schedule from scratch).
    \item \textbf{Latency cost is dominated by P2.} The ${\approx}2{\times}$ slowdown over the RT-DETR baseline ($53.3 \to 25.9$ FPS) is essentially entirely from adding the stride-4 feature level. The feedback module itself, especially with the level mask, is a small fraction of inference cost.

    We see three concrete paths to recover the speed without losing the small-object gain. They are not exclusive --- a final system could combine them.

    \begin{itemize}
        \item[(i)] \emph{Train with P2, infer without.} Use P2 during training to improve small-object learning, then remove it at inference. \textbf{Goal:} keep most of the $+0.99$\,$\mathrm{AP}_S$ gain without the P2 computational cost. This is the highest-leverage experiment: it directly tests whether P2 is doing the work, or whether the feedback path alone is.
        \item[(ii)] \emph{Lighter P2 path.} Reduce the cost of the stride-4 branch by using fewer channels in the P2 projection or by replacing the full input-projection block with a single $1{\times}1$ convolution. \textbf{Goal:} preserve small-object improvements while reducing the slowdown caused by the high-resolution P2 feature map.
        \item[(iii)] \emph{Knowledge distillation.} Distil the small-object improvements of the full v2 model into a faster baseline RT-DETR student network restricted to P3--P5. \textbf{Goal:} recover baseline inference speed while retaining the $\mathrm{AP}_S$ gains learned from P2 feedback during training.
    \end{itemize}
    \item \textbf{COCO only.} Domains where small objects dominate (aerial imagery, traffic surveillance) might show a larger or smaller gain; we did not test there.
\end{enumerate}
